---
title: "Inury_assignment"
output: html_document
---
```{r}

library(readr)
library(dplyr)

injury <- read_csv("~/Documents/mxb344/MAB624_Sem 2 2015_THOMPSON/assessment/assignment_4/injuries.csv")
injury <- 
  injury %>% mutate(Days = Hours/8)

naive_fit <- glm(Injuries ~ Experience + Safety + offset(log(Days)),
                 family = poisson(link="log"),
                 data=injury)
summary(naive_fit)
```

# Diagnostics
### Pearson Residual
```{r}
test_stat <- sum(residuals(naive_fit, type="pearson")^2)
test_stat
```
```{r}
pchisq(test_stat, naive_fit$df.residual, lower.tail = F)
```
Fits Really Well!

### Plots
```{r}
plot(naive_fit)
plot(naive_fit$fitted.values, residuals(naive_fit,type="deviance"))
plot(naive_fit$fitted.values, naive_fit$residuals)
```


#Interpretation
##Relative risk
#Get the unit right for poisson!
#Use crime example.
No std process increases injury rate by exp(.20985)
```{r}
exp(.20985) 
```
That's 23% increase in rate of injury

#How many more hours can an inexperienced worker (< 1) who works in the certifcation and review framework
#expect to work before injury than a worker who works in a regime with no standardised process for equipment training?
```{r}
r_CR <- exp(-5.71159 + .04776)
r_no_std <- exp(-5.71159 + .20985)
```

### Dispersion
```{r}
disp_fit <- glm(Deaths ~ Age + Smoke + offset(log(Pop)),
                family = quasipoisson(link="log"),
                data=smoker)
summary(disp_fit)
```


