---
title: "R Notebook"
output: html_notebook
---

#Modifications to data set
```{r}
library(readr)
library(dplyr)

crime_data <- read_csv('~/Documents/mxb344/MAB624_Sem 2 2015_THOMPSON/assessment/assignment_2/punishment.csv')

#Exploratory analysis


crime_data_mod <-
  crime_data %>% 
  mutate(total_crimes = Pop*Crime) %>%
  select(-Crime)

pairs(crime_data_mod %>% mutate(log_crime = log(total_crimes)))
# Needs to be linear on the scale of log(crime)!
```
Looks Like pop is related super-linearly with population
```{r}
crime_data_mod <- 
  crime_data_mod %>% mutate(Pop_sq = Pop^2)
```


#Proprocessing check for correlation
```{r}
library(caret)
caret::findCorrelation(cor(crime_data_mod %>% select(-total_crimes)), names = T)

```
Sugegsts ignore Po1, it is highly correlated with Po2.

#Naive Poisson fit
```{r}
crime_fit <- glm(family = poisson(link="log"),
                 data = crime_data_mod,
                 formula = total_crimes ~ .
                 )

summary(crime_fit)
```
## Diagnostics
### Test of Scaled Residual Deviance
#### Pearson Residual Deviance
```{r}
pearson_resids <- ((crime_fit$y - crime_fit$fitted.values)/sqrt(crime_fit$fitted.values))^2 %>% sum()
pearson_resids == resid(crime_fit, type="pearson")^2 %>% sum()
resid(crime_fit, type="pearson")^2 %>% sum()
```
It's approximately equal to residual deviance... well same ballpark.

Try Chisq test:
```{r}
pchisq(pearson_resids, df = crime_fit$df.residual, lower.tail = F)
```
It seems like this suggests a terrible fit to data.

#### Scaled Residual Deviance 
Deviance Residual is scaled by dispersion, Phi.
- for introducing deviance there is a sweet normal cancellation to sums of squares GLM book p24.
- GLM boo p 33 for exponential family deviance/scaled deviance.
  - each observation's contribution to the deviance? Deviance residual.
- Calculation of Poisson deviance residual
```{r}
dev_resids <- sqrt(((crime_fit$y * log(crime_fit$y/crime_fit$fitted.values)) - (crime_fit$y - crime_fit$fitted.values)) * 2 )* sign(crime_fit$y - crime_fit$fitted.values)
resid(crime_fit, type="deviance")^2 %>% sum() == sum(dev_resids^2)
resid(crime_fit, type="deviance")^2 %>% sum()
```
Checks out with reported residual deviance above.

Plot this:
```{r}
plot(x=crime_fit$fitted.values, y= dev_resids)
```

####VIF
```{r}
library(car)
#multi-collinearity
vif(crime_fit)
```
This is the generalised VIF.




#Offset Poisson Fit
The offset idea is wrong... it's the number of criminals it's the number of offenses.
```{r}
offset_fit <- crime_fit <- glm(family = poisson(link="log"),
                 data = crime_data_mod,
                 formula = total_crimes ~ M  + Ed + Po2 + LF + M.F + NW + 
                   U1 + U2 + Wealth + Wealth^2 + Wealth*Po2 + Prob*Po2 + Wealth*Ineq + Ineq + Prob + Time + offset(log(Pop*100000))
                 )

summary(offset_fit)
#Removed Po1 because it was not significant.
```
## Diagnostics
### Test of Scaled Residual Deviance
#### Pearson Residual Deviance
```{r}
pchisq(sum(resid(offset_fit, type="pearson")^2) , df = offset_fit$df.residual, lower.tail = FALSE)
```
#### Plot Deviance Residuals
```{r}
plot(x=offset_fit$fitted.values, y=resid(offset_fit, type="deviance"))
```

### Check for free dispersion
```{r}
qoffset_fit <- crime_fit <- glm(family = quasipoisson(link="log"),
                 data = crime_data_mod,
                 formula = total_crimes ~ M + So + Ed + Po2 + LF + M.F + NW + 
                   U1 + U2 + Wealth + Ineq + Prob + Time + offset(log(Pop*100000))
                 )
summary(qoffset_fit)
```
Still looking pretty shit.

# Tuned Model
```{r}
crime_fit_tune <- glm(family = poisson(link="log"),
                 data = crime_data_mod,
                 formula = total_crimes ~ M + So + Ed + Po2 +
                   LF + M.F + Pop + I(Pop^2) + NW + I(NW^2) + U1 + U2 + 
                   Wealth + Prob + Time
                 )

summary(crime_fit_tune)
```
See if step can help
```{r}
library(MASS)
step_obj <- stepAIC(object = crime_fit_tune, scope = total_crimes ~ .^2)


plot(step_obj)
summary(step_obj)
```
## Check diagnostics
### Dispersion
```{r}
step_obj$call
glm(formula = total_crimes ~ M + So + Ed + Po2 + LF + M.F + Pop + 
    I(Pop^2) + NW + I(NW^2) + U1 + U2 + Wealth + Prob + Time + 
    M.F:Time + M.F:NW + So:NW + Pop:I(Pop^2) + Prob:Time + M.F:Pop + 
    So:M.F + So:Ed + Pop:Prob + M:So + Po2:Wealth + U1:U2 + M.F:I(NW^2) + 
    Po2:Prob + LF:M.F + LF:U2 + M:U2 + U1:Time + Wealth:Time + 
    LF:Wealth + Ed:Po2 + NW:U2 + M:Pop + LF:I(NW^2) + So:U2 + 
    I(Pop^2):I(NW^2) + LF:NW + Po2:I(NW^2), family = quasipoisson(link = "log"), 
    data = crime_data_mod) %>% summary()
```
A bit under dispersed.
### Pearson Corellation
```{r}
pchisq( sum(residuals(step_obj, type="pearson")^2) , df = step_obj$df.residual   )
```

#Limited Step Model
```{r}

lm_mod %>% summary()

step(lm(log(total_crimes) ~ I(Pop^2), data=crime_data_mod), scope = log(total_crimes) ~ M + So + Ed + Po2 +
                   LF + M.F + Pop + I(Pop^2) + NW + I(NW^2) + U1 + U2 + 
                   Wealth + Prob + Time) %>% summary()
```
So it turns out a plain old linear model fits this data just great.
